---
title: "Meow"
author: "Sarah Wood"
date: "12/15/2021"
output: 
  html_document: 
    theme: journal
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
getwd()
setwd("C:/Users/sarah/OneDrive/Documents/R/Fall2021/EcologyProject")
getwd()
```

------------------------------------------------------------------------

------------------------------------------------------------------------

# [SETUP]{.ul}

### **Libraries**

> `library(dplyr)`
>
> `library(knitr)`
>
> `library(tidyr)`
>
> `library(ggplot2)`
>
> `library(forcats)`
>
> `library(inspectdf)`

```{r libraries, include=FALSE}
library(dplyr)
library(knitr)
library(tidyr)
library(ggplot2)
library(forcats)
library(inspectdf)
library(rmarkdown)
```

 

###   **Options for** `str()` **for narrow transcript windows**

```{r str() options, results = 'hide'}
#setting up str function to a narrow transcript window setting
options(width = 100, str = strOptions(strict.width = "cut"))
str(lsfit(1:9,1:9))
```

 

 

### **Compact Checking Function**

> Making a function that can be applied to an object and....
>
> 1.  Sense if it's a column, row, or data frame
>
> 2.  Applies certain checking functions depending on the result of the previous function
>
>     1.  If a `data frame` --> `str`, `colnames`, `rownames`, `dim`, and `class`
>
>     2.  If a `row`or `col` --> `table`, `names`, `length`, `class`, and `unique`
>
> 3.  Returns the outputs all at once

##### **Establishing the functions**

```{r compact checking function1}
#making a fnction for checking dataframes
check_dfsw <- function(x) {
  str(x)
  
  colnames <- colnames(x)
  rownames <- rownames(x)
  dims <- dim(x)
  class <- class(x)
  
  output <- list("column names" = colnames, #making a list of outputs bc return() only takes one item
                 "row names" = rownames, "dimensions" = dims, "class type" = class)
  return(output) 
}

#making a function for checking rows or columns or lists
check_listsw <- function(x) {
  
  table <- table(x)
  names <- names(x)
  length <- length(x)
  class <- class(x)
  unique <- unique(x)
  
  output <- list("table" = table, "names" = names, "length" = length, 
                 "class type" = class, "unique values" = unique)
  
  return(output) 
}
```

 

##### **Making a function that that checks data type then does the previous functions**

Based on this:

> `dim(data)`
>
> `dim(data$paper_ID)`
>
> `dim(data[1,])`

```{r compact checking function2}
#the function
checking <- function(x){
  dimsum <- dim(x)      #determines the dimensions of x and assigns it to dimsum
  ifelse (((is.null(dimsum) == FALSE) & ((dimsum[1]==1) == FALSE)),   
          #things that are df only if dimsum has is NOT null
          # and dimsum's first value (the # of rows) does NOT equal 1 
          #(meaning not a single row)
          
          results <- check_dfsw(x), 
          #if the above is true = x is df = will do the df checks 
          #and store results in "results"
          results <- check_listsw(x)) 
          #if the above is false = x is row = will do list checks 
          #and store results in "results"
  
  ifelse ((is.null(dimsum) == TRUE), 
          #meaning: if the dimensions are null = if it's a column
          results <- check_listsw(x), 
          #will do the list checks and store results in "results"
          NA)                         
          #will do nothing if not a list 
  return(results) #prints the results
}
```

 

 

### **Data Loading**

```{r Data Loading, results='hide'}
data <- read.csv("ecol2021_reconfigured_data.csv", 
                 sep = ",", 
                 as.is = TRUE, 
                 na.strings = "NA", 
                 row.names = 1, 
                 fill = TRUE, 
                 skip = 0) 
```

 

 

 

# [**DATA FORMATTING**]{.ul}

### **Changing Rownames**

> `data$ID` This column has the IDs of each model. we'll make R use these as row names.
>
> 1.  Adding "ID" to all the entries in the `data$ID` column
>
> 2.  Assigning the values in the `ID` column to the rownames of `data`
>
> 3.  `checking(data)` 

```{r changing rownames, echo=TRUE}

head(rownames(data))

data$ID <- sub("^", "ID", data$ID) #adding "ID" before all entries in data$ID
rownames(data) <- data$ID #assigning the vaues in the ID column to the rownames

head(rownames(data)) #check. it worked!
```

### **Cutting off unnecessary columns**

> 1.  Finding which ones we need to remove using `colnames(data)`
>
> 2.  Shaving off some unnecessary stuff by replacing the dataset with one that...
>
>     -   Exlcudes the `$totals` row (row 563)
>
>     -   Retains all columns except ones named `$ID`
>
> 3.  Using `dplyr` to select just the columns we want to keep and replace the dataframe with one that just has those columns

```{r cutting the fat, echo=TRUE}

#1
#colnames(data)

#2
data <- data[1:562,] 
data <- data[,!names(data) == "ID"] 

head(colnames(data)) #checking
tail(rownames(data))

#3
colnames(data)
#doing it in a copy just so we get it right before replacing the dataframe
data2 <- data %>% select(c(1,3,6,8:10,         #types and IDs
                          17:21,32,35,36,45,  #non crop explanatories
                          51,54:60,62,63,     #border-related expl.
                          67:70,              #forest border-related
                          71:73))             #signif stuff that is misplaced
    
# data <- select(data, c(1:8, 17:73)) is another way
colnames(data2)  #checking 
rm(data2) #removing because we don't need it anymore

#replacing the dataframe
data <- data %>% select(c(1,3,6,8:10,         #types and IDs
                          17:21,32,35,36,45,  #non crop explanatories
                          51,54:60,62,63,     #border-related expl.
                          67:70,              #forest border-related
                          71:73)) 
```

|                  Group |   Model Characteristic    |             `colname`             | `class` |
|-----------------------:|:-------------------------:|:---------------------------------:|---------|
|             Model Type |         Paper ID          |            `paper_ID`             |         |
|             Model Type |           Model           |           `model_type`            |         |
|          Response Type | Generalized Response Type |      `resp_type_generalized`      |         |
|          Response Type |       Result Value        |              `value`              |         |
|          Response Type |     Result Value type     |           `value_type`            |         |
|               Strength |        Significant        |             `signif`              |         |
|               Strength |       "Best" Model        |          `value_isBEST`           |         |
|               Strength |   "Best" Status Support   |       `value_best_num_var`        |         |
|       Explanatory Type |   Explanatory Variables   |           `exp_num_var`           |         |
| Explanatory Type^land^ |     Border type - ANY     |        `exp_isBORDERTYPE`         |         |
| Explanatory Type^land^ |  Border type - Buckthorn  |      `exp_border_buckthorn`       |         |
| Explanatory Type^land^ |    Border type - Crops    |     `exp_isBORDERTYPE_crops`      |         |
| Explanatory Type^land^ |     (crops - alfalfa)     |     `exp_bordertype_alfalfa`      |         |
| Explanatory Type^land^ |      (crops - corn)       |       `exp_bordertype_corn`       |         |
| Explanatory Type^land^ |     (crops - canola)      |      `exp_bordertype_canola`      |         |
| Explanatory Type^land^ |     (crops - cereals)     |     `exp_bordertype_cereals`      |         |
| Explanatory Type^land^ |     (crops - soybean)     |     `exp_bordertype_soybean`      |         |
| Explanatory Type^land^ |   (crops - generalized)   | `exp_bordertype_generalizedcrops` |         |
| Explanatory Type^land^ |   Border type - Grassy    |     `exp_isBORDERTYPE_grass`      |         |
| Explanatory Type^land^ |   Border type - Forest    |     `exp_isBORDERTYPE_forest`     |         |
| Explanatory Type^land^ |    (forest - distance)    |     `exp_isBORDERTYPE_forest`     |         |
| Explanatory Type^land^ |   (forest - proportion)   | `exp_bordertype_forestproportion` |         |
| Explanatory Type^land^ |   (forest - diversity)    | `exp_bordertype_forestdiversity`  |         |
| Explanatory Type^land^ |    (forest - distance)    |  `exp_bordertype_forestdistance`  |         |
| Explanatory Type^land^ |     Field Parameters      |           `exp_isFIELD`           |         |
|       Explanatory Type |           Scale           |           `exp_isSCALE`           |         |
|       Explanatory Type |           Time            |           `exp_isTIME`            |         |
|       Explanatory Type |            SBA            |            `exp_isSBA`            |         |
|       Explanatory Type |       SBA Predators       |           `exp_isPRED`            |         |
|       Explanatory Type |        Null or NA         |            `exp_isNA`             |         |
|       Explanatory Type |           Other           |           `exp_isOTHER`           |         |

```{r cutting the fat}
    


#gonna relocate the columns at the end to the beginning
    data <- relocate(data, c("value_isBEST", "value_best_num_var", "signif"), .before = exp_isNA)
    data <- relocate(data, value_best_num_var, .before = value_isBEST)
    #^ this is dplyr again
    checking(data) #sexy. cols 11 through 65 are now just binary


    ############################################################
    ############ changing to factors ##################### #####
    
    data[sapply(data, is.integer)] <-                       #sapply is like lapply below
        lapply(data[sapply(data, is.integer)], as.factor)   #taking data that is numeric and factorizing it
            #lapply returns a list of same length of x, each element of which is 
            #the result of applying FUNCTION to the corresponding element of x
    checking(data) #worked
    
    #doing the same for certain categorical variables 
    
    data[,1:4] <- lapply(data[,1:4], as.factor)  #making columns 1-4 factors
    data[,7] <- as.factor(data[,7])              #making column 7 a factor
    checking(data) #worked
    

    str.data <- as_tibble(str(data))
```

```{r, rearranging}
data <- relocate(data, c("value_isBEST", "value_best_num_var", "signif"), .before = exp_isNA)
    data <- relocate(data, value_best_num_var, .before = value_isBEST)
    #^ this is dplyr again
    checking(data) #sexy. cols 11 through 65 are now just binary
>
```

```{r}

```

 

## Subsetting

You can also embed plots, for example:

```{r pressure, echo=FALSE}

##################    SUBSETTING     #################################### ######
    
    #just the categorical variables 
    checking(data)
    datacat <- data[ , c(1:4,7:64)]
    checking(datacat)
    
    #just the bigger groups of variables (with noncat)
    checking(data)
    datamain <- data[ , c(1:17,26,29,30,39,45,47:49,55,57,61)]
    checking(datamain)
    
    # numerics 
    datanum <- data[ , c("value", "exp_num_var", "value_best_num_var")]
    datanum[,1:3] <- lapply(datanum[,1:3], as.numeric)
    checking(datanum)
    
    #data.exp.other.m  = data for other explanatory variables to be in a matrix
    data.exp.other.m <- data[ , c("exp_isBORDERTYPE", "exp_isNA", "exp_isTIME",
                                  "exp_isSCALE", "exp_isPRED", "exp_isFIELD", 
                                  "exp_isSBA", "exp_isOTHER",
                                  
                                  "signif", "value_isBEST", "value_best_num_var", 
                                  "exp_num_var", "paper_ID", "resp_type_generalized"
                              )]
    checking(data.exp.other.m)
    
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}
################################################################################
##################    NOTES          #################################### ######
  
    ############################################################
    ################### Notes on variables ############### #####   
    
      #explanatory (chr)
      #value      (num)
    
      ###################### ID ones  (factors)
      
      #paper_ID
      #model_type
      #year
      #resp_type_generalized
      #value_type
      
      ###################### strength ones (factors)
      
      #signif
      #value_isBEST
      #value_best_num_var 
      
      ###################### overall explanatory ones (factors)
      
      #exp_num_var    
      
      # isNA    isTIME    isSCALE   
      
      # isOTHER   SOYPRED   FUNGUS    isSBA   isPRED    isPREDTYPE    isFIELD
      # isBORDERTYPE    
      # border_buckthorn    isBORDERTPE_crops   isBORDERTYPE_grass    isBORDERTYPE_forest
      
      
      ####################### explanatory ones delineated (factors)
      
      # isOTHER  
          # plantnutrition  leaf_percK    leaf_perc_N   precipitation   growingdegreedays   
          # functdiversity  aspect        intercept
      # SOYPRED   
      # FUNGUS    
      # isSBA 
          # captured_SBA    SBA_abundance
      # isPRED    
          # pred_migration   pred_withheld   pred_richness   pred_abundance
          # predtype_lacewings   predtype_spiders    coccinellidae   parasitoids
      # isPREDTYPE   
          #p redtype_lacewings   predtype_spiders    coccinellidae   parasitoids
      # isFIELD
          # fieldID   fieldperimeter    fieldarea   fieldwidth    fieldtillage_notill
      
      
      # isBORDERTYPE   
          # border    landscape_complexity    bordertype_richness
      # border_buckthorn    
      #isBORDERTPE_crops   
          # bordertype_alfalfa    bordertype_corn   bordertype_canola   
          # bordertype_cereals   bordertype_soybean   bordertype_generalizedcrops
      # isBORDERTYPE_grass  
          # bordertype_grass    bordertype_prairie    bordertype_pasture
      # isBORDERTYPE_forest
          # bordertype_forestdistance   bordertype_forestproportion   forestdiversity
    
      
```
